NHL realignment effects on the conference gap

### 2020-12-14 Paper.Rmd
upload my part
uploaded my part and fixed a lot of stuff

---
title: "NHL realignment effects on conference gap"
author: "Kilbourne Charrier, Michael Bauers, Clinton, Sam Downs"
output: pdf_document
---

@@

# Motivation

## Introduction

>>>>

---
title: "NHL realignment effects on the conference gap"
author: "Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs"
output: pdf_document
---

@@

# Motivation
   In the 2013-14 National Hockey League (NHL) season, the league made some changes to the conference numbers. The Winnipeg Jets moved from the Eastern conference to the Western conference. This created an imbalance in the conference numbers giving the Eastern conference sixteen teams and leaving the Western conference with only fourteen teams. From the start, it already seemed unfair to have more teams in one conference than the other. If every team had equal skill, the fourteen teams in the western conference had an equal one four in seven chance of making it to the playoffs. In contrast, the sixteen teams in the eastern conference had a fifty percent chance of making it to the playoffs. This uneven probabillity will be known for the rest of the paper as the conference gap.

  Now how could an organization as big as the NHL allow for that kind of discrepency between the conferences? As a continuation of our research we wanted to look at another season with a similiar balance shift in the conferences to see how that change could have affected the outcome of the season. The season we chose was the recent expansion draft season of 2017-18. Before the start of the season, the NHL added a new franchise in the Las Vegas Golden Knights. This team was set to join the Western conference. This shifted the conference split from fourteen to sixtten West to East to a closer fifteen to sixteen split. We wanted to see if our result was statistically significant from both the conference gap of an even season and the conference gap from our tested 2013-14 season.

## Introduction
  In 2011 when the Atlanta Thrashers became the Winnipeg Jets. A problem was presented as Winnipeg is considerably far from the east coast of the North American continent. Because of this and how the season is set up, the average travel time for teams playing winnipeg were fairly large as eastern conference teams play each other more often. Just before the start of the 2013-14 season, the league offices for the NHL decided to move over the Winnipeg Jets from the Eastern conference to the Western conference and in return the Eastern conference aquired the Columbus Blue Jackets and the Detroit Redwings. This left the the two conferences at an uneven balance. The Eastern conference now had sixteen teams while the Western conference only had fourteen teams. This presented a curious problem as for both conferences, eight teams make the playoff brackets. It is easy to see the problem at hand that the western conference has just as many playoff slots for having two less teams. The goal of our research and the paer we leaned on, was to discover what that difference could mean in the grander sense of the league moving forward with this set up. We did this based on the average points and wins necessary to make the playoffs for both conferences.

  To further our study, we also looked at the 2017-18 season. This season is important becasue it gave the introduction of a new team in the league none as the Las Vegas Golden Knights. This team was added to the Western conference makiing the split slightly more even. Our goal for this was to see how adding a single team could change the skill outcome for the playoffs. We looked at the same measures to understand the new conference gap as our previous analysis of the 2013-14 season.
